### 2018-06-21 source/Ch7-3C295/3c295.tex
finished foreword, section 6, catalog annex

\chapter{Imaging 3C295 with International LOFAR}\label{section.3c295}
\minitoc
% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}
\pg
Our aim in this section is to create a high-resolution (0.1$''$) model of 3C295, which did not exist when this project began for LOFAR frequencies. This means creating a model that will allow us to find good phase-calibration solutions for LOFAR international baselines, and thus use the international stations to create a high-resolution image of the EGS.
%While we also solve for amplitude gains, we know that we will need to correct the total source flux in each frequency subband as our initial spectral model is not necessarily correct.

\pg
One major constraint in this project is that 3C295 is fully resolved at these resolutions, which makes deconvolution difficult. As such, we consistently use a $uv$-cut of $100$km throughout our imaging in this section, to limit the impact of diffuse emission (which is very hard to deconvolve properly, and introduces artefacts in the image when improperly done) on our images. This choice of $uv$-cut is a compromise between removing diffuse emission in the images and providing sufficiently good conditioning to the calibration inverse problem\footnote{The more uncorrupted (i.e. flagged for RFI) visibilities are available, the better the estimation of gain solutions - see \citetads{2015MNRAS.449.2668S} for the demonstration.}. Provided that the phases of short-baseline gains do not evolve over the self-calibration procedure, then the information on flux structure associated to these spatial fringes is not lost.

@@

\pg
As outlined in \cref{section.RIME}, calibration is the process of solving for Jones matrices, which are associated to individual antennas. Here, the only source in our model (the source that dominates, by far, on the total emission in the field) is 3C295; we can thus limit ourselves to \textit{direction-independent} calibration, which is equivalent to performing direction-dependent calibration in the direction of 3C295 only. The associated RIME can be written as:
\begin{align}
\Vis_{pq}^{t,\nu} &= \Gjones_p^{t,\nu} \left( \sum_{s=3C295} \Ejones_{sp}^{t,\nu} \Kjones_{sp}^{t,\nu} \Bmatrix \left(\Kjones_{sq}^{t,\nu}\right)^H \left(\Ejones_{sq}^{t,\nu}\right)^H \right) \left(\Gjones_q^{t,\nu}\right)^H\\
		  &= \Gjones_p^{t,\nu} \Ejones_{3C295,p}^{t,\nu} \Bmatrix_{3C295}^{t,\nu} \left(\Ejones_{3C295,q}^{t,\nu}\right)^H \left(\Gjones_q^{t,\nu}\right)^H\\
		  &= \Jones_{3C295,p}^{t,\nu} \Bmatrix_{3C295}^{t,\nu} \left(\Jones_{3C295,q}^{t,\nu}\right)^H
\end{align}
where we explicitly model the LOFAR beam during both calibration and imaging, so as to ensure that the true sensitivity of each baseline to this out-of-phase-centre calibrator is taken into account appropriately.

@@

\pg
We begin by using a simple spectral index of $\alpha=-0.8$, which is typical of synchrotron sources at these frequencies. It is not strictly correct, but gives a good first-order approximation of the integrated flux of 3C295 as a function of frequency. In so doing, we introduce a frequency- and baseline-dependent error factor:
\begin{align}
\hatJones_{3C295,p}^{t,\nu} &\approx \alpha_{pq}^\nu \Jones_{3C295,p}^{t,\nu}
\end{align}
where $\hatJones_{3C295,p}^{t,\nu}$ is our estimation of the Jones matrix $\Jones_{3C295,p}^{t,\nu}$ for time $t$ and frequency $\nu$, and the baseline-dependence of $\alpha_{pq}^\nu$ arises because 3C295 is not at phase centre. If it was, then $\alpha$ would be a function of frequency alone. $\alpha$ varies smoothly as a function of frequency. 
Because we solve for the full Jones matrix, we cannot apply phase-only corrections over multiple self-calibration rounds, as is standard practice. Over self-calibration rounds, we expect that $\alpha$ will change, but that it will remain smooth in frequency. The integrated flux of 3C295 over different passes of self-calibration is shown in \cref{fig.SC.ampfluct}: as this figure shows, the model amplitude can diverge significantly over the self-calibration process, sometimes even moving closer to the inital integrated flux, but often moving farther.

% images:
% IMAGES/3c295.fromVLA.6sb.withbeam.pass[1-5].int.model.fits IMAGES/3c295.fromVLA.60sb.withbeam.pass[1-3].int.model.fits IMAGES/MSlist.fullBW.LOBOS6.MSlist.fullBW.txt.uniform.int.model.fits
% script: /home/offler/plotSCamp.py
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/AmpFluct.png}
\caption{\label{fig.SC.ampfluct} Amplitude fluctuation of the models resulting from self-calibration iterations, expressed as a ratio of the model's integrated flux divided by the Scaife \& Healde integrated flux model at the appropriate frequency. Note that the integrated model flux for 324sb-pass1 (the penultimate data in this plot) was lost to a careless command: the value shown here is just the average of the two neighbouring values.}
\end{figure}

\pg
Because $\alpha$ is smooth in frequency, we can apply a frequency-dependent amplitude correction (i.e. we can ``bootstrap") by relying on the work of Scaife \& Heald \citep[see][]{arse}. This correction will be done in the imaging plane, since $alpha$ is a function of $uv$-position (but not sky position). This is, again, because 3C295 does not lie at phase centre: if it did, the correction could be done equivalently in visibility or image space. By performing it in the image-plane, we rid ourselves of the $uv$-dependence. The resulting signal-to-noise improvement is shown in \cref{fig.SCrms}. It is calculated on the bootstrapped residual images resulting from each round of the self-calibration process - as such, the only thing that changes is the noise, since the signal is corrected to be the appropriate constant As we can see, the noise falls nicely with self-calibration passes. Tragically, the results from the first full-bandwidth calibration run with the 60-subband model were lost before this plot could be made, and so only the results from the second full-bandwidth self-calibration pass remain. We can see that it is a small improvement over the 60-subband model. 
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/SCrms.png}
\caption{\label{fig.SCrms} Bootstrapped residual RMS for each calibration pass. Note that the data for 324sb-pass1 (the penultimate data in this plot) were lost to a careless command: the value shown here is just the average of the two neighbouring values.}
\end{figure}

\pg
The large decrease in RMS between the last 6-subband residual image and the first 60-subband residual image in \cref{fig.SCrms} is significant: we would expect a similar jump when going from 60 to 324 subbands. The fact that we do not is a strong indicator that we are digging into the limits of what our strategy can achieve by the end of the second full-bandwidth self-calibration pass. This is likely due to the fact that our initial sky model (and all subsequent models) only include 3C295: the fact that all the other sources in the primary beam are not included in the model is bound to introduce bias our gain solutions and artefacts in our images. This is a likely explanation for this behaviour.

@@

\pg
Because no high-resolution models exist for 3C295 at our observing frequencies, which span quite a large bandwidth, we start by making a model from a high-resolution image taken at 8.7GHz, shown in \cref{fig.vla.3c295}. We start with a model which is as close to the expected ``true" brightness distribution because the self-calibration problem is not convex: while the calibration inverse problem itself is convex (i.e. we expect to converge towards the ``true" gains regardless of the initial conditions, given enough SNR), the degeneracy introduced by imaging in the self-calibration procedure means that one could well get stuck in a local minimum. An exhaustive method would be to perform a full Monte Carlo analysis, testing every possible flux distribution for every pixel at every frequency to ensure that we do, in fact, find the model and gains which give the global minimum fit, rather than a local minimum fit. This would be extremely time-consuming, however, and falls outside the scope of this work. We therefore simply try to start as close as possible to where we expect the global minimum to fall.

\pg
We acquire the initial calibration model by extracting features from a NASA/IPAC Extragalactic Database\footnote{\hyperref[here]{https://ned.ipac.caltech.edu/}} image of 3C295 \citepads[see][]{1991AJ....101.1623P}, shown in Fig. \ref{fig.vla.3c295}. This feature extraction is carried out using PyBDSM \citepads{2015ascl.soft02007M}, changing all extracted Gaussians into points\footnote{This practice ensures that wrongly-estimated Gaussians do not end up introducing unphysical bias during calibration.}.
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/{3C_295_I_8.7GHz_lbs2003.fits}.png}
\caption{\label{fig.vla.3c295}VLA observation of 3C295 at 8.7 GHz. Pixel size is $0.2''$.}
\end{figure}
% command: python ~/Scripts/PngCutoutFromFits.py --filename 3C_295_I_8.7GHz_lbs2003.fits --RA 14:09:33.418 --Dec 52:26:13 --size .12 --vmin -0.5 --vmax 4 --invert
\pg
Because we apply a $uv$-cut of 1km when calibrating, and a $uv$-cut of 100km when imaging, we need to make sure that we do not introduce spurious structure in the core and remote station gains. Similarly, we can check the quality of our self-calibration by looking at the structure of the international station gains. Gain structure can consist both of genuine atmospheric \& instrumental effects, but can also consist of unmodeled flux being absorbed into the gain solutions. At first, this is unavoidable: if flux is resolved by a baseline but not included in the model, then it will be absorbed into the gains. As we improve our model, we expect the gain curves to flatten and become smoother. If this happens for our international station gains, then we have a strong indication that we are in fact recovering physical signal rather than adding artefacts into our model. Similarly, we expect core and remote station gains to remain mostly static: if this is the case, then our strategy does not bias our gain solutions for these stations away from the underlying truth.

\pg
Let us begin by checking whether our core and remote station gains change over time due to our self-calibration procedure. Here, we look at the gains of SB155, on which we applied the initial and final gain solutions. This subband was included in the 60-subband self-calibration step, but not the 6-subband self-calibration step.
\begin{figure}[h!]
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.firstpass}.png}
\caption{\label{fig.sb155.gains.cs201.before} CS005, first calibration}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.lastpass}.png}
\caption{\label{fig.sb155.gains.cs201.after} CS005, last calibration}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.RS305.before} RS305, first calibration}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.RS305.afer} RS305, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.CR} Gain amplitude \& phase for CS201 of SB155}
\end{figure}

\pg
As we can see from \cref{fig.sb155.gains.CR}, there is little change in amplitude structure between the first pass of calibration (with the model extracted from \cref{fig.vla.3c295}) and the last pass of calibration (after the successive self-calibration passes). Similarly, there is little apparent change in phase structure (structure which is difficult to see, as the phase wraps from $-\pi$ to $\pi$). It is important to note, however, that there is a difference in the \textit{level} of the amplitude curve: in the first pass of calibration, the average amplitude value is generally higher than in the last pass of calibration. This is expected behaviour, as there is less diffuse emission in the $uv$-cut imaging model of 3C295 than in the initial calibration model. This is equivalent to applying a choice of weighting scheme which down-weighs diffuse emission (i.e. a form of uniform weighting).

\pg
As for the international station gains, they are shown below for three of the seven international stations used for our observation, in \cref{fig.sb155.gains.international}. We show only one of the four German stations. As we can see, while structure very clearly remains in some of these amplitude curves (notable in the last two hours \cref{fig.sb155.gains.DE602.after}), some of the structure disappears from these amplitude curves. It should be noted that the last two hours of the observation are generally much noisier than the first 6; the last quarter of these plots show this. Quality-based weighting schemes compensate for this effect, and minimise the associated image deterioration. Note that the phase wraps extremely fast in these plots: this is because it is given with reference to CS001HBA0, the first antenna in our array. Because the international stations are much farther away from this core station than other core or remote stations are, we expect the associated phase solutions to wrap faster.
\begin{figure}[h!]
%\begin{subfigure}{.49\\textwidth}
%\includegraphics[width=\linewidth]{images/{SB155.DE601HBA.firstpass}.png}
%\caption{\label{fig.sb155.gains.DE601.before} DE601, first calibration}
%\end{subfigure}
%\begin{subfigure}{.49\\textwidth}
%\includegraphics[width=\linewidth]{images/{SB155.DE601HBA.lastpass}.png}
%\caption{\label{fig.sb155.gains.DE601.after} DE601, last calibration}
%\end{subfigure}
\begin{subfigure}{.49\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.DE602.before} DE602, first calibration}
\end{subfigure}
\begin{subfigure}{.49\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.DE602.after} DE602, last calibration}
\end{subfigure}
%\begin{subfigure}{.49\textwidth}
%\includegraphics[width=\linewidth]{images/{SB155.DE603HBA.firstpass}.png}
%\caption{\label{fig.sb155.gains.DE603.before} DE603, first calibration}
%\end{subfigure}
%\begin{subfigure}{.49\textwidth}
%\includegraphics[width=\linewidth]{images/{SB155.DE603HBA.lastpass}.png}
%\caption{\label{fig.sb155.gains.DE603.after} DE603, last calibration}
%\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.SE607HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.SE607.before} SE607, first calibration}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.SE607HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.SE607.after} SE607, last calibration}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.UK608HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.UK608.before} UK608, first calibration}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.UK608HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.UK608.after} UK608, last calibration}
\end{subfigure}

\caption{\label{fig.sb155.gains.international} Gain amplitude \& phase for international stations of SB155}
\end{figure}

\pg
So far, we have shown the gain solutions for one subband, out of a total of 324 used subbands. The spectral behaviour of the amplitude and phase solutions can give us insight into whether we are absorbing true flux or calibration artefacts into our model: if our calibration model becomes more accurate to the true underlying flux, we expect e.g. the phase of core station calibration solutions to remain the same (as we are improving high-resolution details of our model) while our international station calibration solutions would become flatter in amplitude. Similarly, we expect to see large-scale frequency-dependent behaviour in the phase solutions of international stations, with smaller fluctuations around this behaviour: the effect of differential TEC, ionospheric turbulence, etc. Indeed, while the core stations (and, to a lesser extent, the remote stations) can be approximated as having the ``same ionosphere" above them, this is not the case for international stations: this is the source of the difficulties in calibrating these stations. As such, for international stations, we expect to see complex structure in the phase gain solutions, characteristic of the differences between the local atmosphere above the core and the international stations. For core stations, we only expect to see structure characteristic of scintillation in the phase gain solutions.

@@

\begin{figure}[h!]
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.gains.CS005.before} CS005HBA0, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.gains.CS005.after} CS005HBA0, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.CS005} Plots showing the amplitude and phase solutions for CS005HBA0 at the start and end of self-calibration, respectively. Some areas are empty: in the case of the first calibration, this is because the solutions were not calculated for all subbands. For the last calibration, this is because some subbands were found to be corrupted, and their solutions are thus not shown here.}
\end{figure}
\begin{figure}[h!]
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{CS201HBA0.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.CS201HBA0.before} CS201HBA0, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.CS201HBA0.after} CS201HBA0, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.CS201HBA0} Plots showing the amplitude and phase solutions for CS201HBA0 at the start and end of self-calibration, respectively. Some areas are empty: in the case of the first calibration, this is because the solutions were not calculated for all subbands. For the last calibration, this is because some subbands were found to be corrupted, and their solutions are thus not shown here.}
\end{figure}

>>>>

\chapter{Imaging 3C295 with International LOFAR}\label{section.3c295}
\minitoc
% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}
\pg
Our aim in this section is to create a high-resolution (0.1$''$) model of 3C295, which did not exist at LOFAR HBA frequencies when this project began. This means creating a model that will allow us to find good phase-calibration solutions for LOFAR international baselines, and thus use the international stations to create a high-resolution image of the EGS.
%While we also solve for amplitude gains, we know that we will need to correct the total source flux in each frequency subband as our initial spectral model is not necessarily correct.

\pg
One major constraint in this project is that 3C295 is fully resolved at these resolutions, which makes deconvolution difficult. As such, we consistently use a $uv$-cut of $100$km throughout our imaging in this section, to limit the impact of diffuse emission (which is very hard to deconvolve properly - poor deconvolution introduces artefacts in the restored image) on our images. This choice of $uv$-cut is a compromise between removing diffuse emission in the images and providing sufficiently good conditioning to the calibration inverse problem\footnote{The more uncorrupted (i.e. flagged for RFI) visibilities are available, the better the estimation of gain solutions - see \citetads{2015MNRAS.449.2668S} for the demonstration.}. Provided that the phases of short-baseline gains do not evolve over the self-calibration procedure, then the information on spatial brightness distribution associated to these spatial fringes is not lost.

@@

\pg
As outlined in \cref{section.RIME}, calibration is the process of solving for Jones matrices, which are associated to individual antennas. Here, the only source in our model (the source that dominates, by far, over other sources in the field) is 3C295; we can thus limit ourselves to \textit{direction-independent} calibration, which is equivalent to performing direction-dependent calibration in the direction of 3C295 only. The associated RIME can be written as:
\begin{align}
\Vis_{pq}^{t,\nu} &= \Gjones_p^{t,\nu} \left( \sum_{s=3C295} \Ejones_{sp}^{t,\nu} \Kjones_{sp}^{t,\nu} \Bmatrix \left(\Kjones_{sq}^{t,\nu}\right)^H \left(\Ejones_{sq}^{t,\nu}\right)^H \right) \left(\Gjones_q^{t,\nu}\right)^H\\
		  &= \Gjones_p^{t,\nu} \Ejones_{3C295,p}^{t,\nu} \Bmatrix_{3C295}^{t,\nu} \left(\Ejones_{3C295,q}^{t,\nu} \Gjones_q^{t,\nu}\right)^H\\
		  &= \Jones_{3C295,p}^{t,\nu} \Bmatrix_{3C295}^{t,\nu} \left(\Jones_{3C295,q}^{t,\nu}\right)^H
\end{align}
where we explicitly model the LOFAR beam during both calibration and imaging, so as to ensure that the true sensitivity of each baseline to this out-of-phase-centre calibrator is taken into account appropriately.

@@

\pg
We begin by using a simple spectral index of $\alpha=-0.8$, which is typical of synchrotron sources at these frequencies. It is not strictly correct, but gives a good first-order approximation of the integrated flux of 3C295 as a function of frequency. In so doing, we introduce a frequency- and baseline-dependent error factor:
\begin{align}
\hatJones_{3C295,p}^{t,\nu} &\approx \alpha_{pq}^\nu \Jones_{3C295,p}^{t,\nu}
\end{align}
where $\hatJones_{3C295,p}^{t,\nu}$ is our estimation of the Jones matrix $\Jones_{3C295,p}^{t,\nu}$ for time $t$ and frequency $\nu$, and the baseline-dependence of $\alpha_{pq}^\nu$ arises because 3C295 is not at phase centre. If it was, then $\alpha$ would be a smoothly-varying function of frequency alone.
Because we solve for the full Jones matrix, we cannot apply phase-only corrections\footnote{Indeed, what is the phase of a $2\times2$ matrix of complex numbers?} over multiple self-calibration rounds, as is standard practice. Over self-calibration rounds, we expect that $\alpha$ will change, but that it will remain smooth in frequency. The integrated flux of 3C295 over different passes of self-calibration is shown in \cref{fig.SC.ampfluct}: as this figure shows, the model amplitude can diverge significantly over the self-calibration process. %, sometimes even moving closer to the inital integrated flux, but often moving farther.

% images:
% IMAGES/3c295.fromVLA.6sb.withbeam.pass[1-5].int.model.fits IMAGES/3c295.fromVLA.60sb.withbeam.pass[1-3].int.model.fits IMAGES/MSlist.fullBW.LOBOS6.MSlist.fullBW.txt.uniform.int.model.fits
% script: /home/offler/plotSCamp.py
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/AmpFluct.png}
\caption{\label{fig.SC.ampfluct} Amplitude fluctuation of the models resulting from self-calibration iterations, expressed as a ratio of the model's integrated flux divided by the Scaife \& Healde integrated flux model at the appropriate frequency. Note that the integrated model flux for 324sb-pass1 (the penultimate data point in this plot) was lost to a careless command: the value shown here is just the average of the two neighbouring values.}
\end{figure}

\pg
Because $\alpha$ is smooth in frequency, we can apply a frequency-dependent amplitude correction (i.e. ``bootstrap") by relying on the work of Scaife \& Heald \citep[see][]{arse}. This correction will be done in the imaging plane, since $alpha$ is a function of $uv$-position (but not sky position). This is, again, because 3C295 does not lie at phase centre: if it did, the correction could be done equivalently in visibility or image space. By performing it in the image-plane, we rid ourselves of the $uv$-dependence. The resulting signal-to-noise improvement is shown in \cref{fig.SCrms}. It is calculated on the bootstrapped residual images resulting from each round of the self-calibration process - as such, the only thing that changes is the noise, since the signal is corrected to be the appropriate constant As we can see, the noise falls nicely with self-calibration passes. Tragically, the results from the first full-bandwidth calibration run with the 60-subband model were lost before this plot could be made, and so only the results from the second full-bandwidth self-calibration pass remain. We can see that it is a small improvement over the 60-subband model. 
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/SCrms.png}
\caption{\label{fig.SCrms} Bootstrapped residual RMS for each calibration pass. Note that the data for 324sb-pass1 (the penultimate data in this plot) were lost to a careless command: the value shown here is just the average of the two neighbouring values.}
\end{figure}

\pg
The large decrease (factor of $\sim2$)in RMS between the last 6-subband residual image and the first 60-subband residual image in \cref{fig.SCrms} is significant: we would expect a similar jump when going from 60 to 324 subbands. The fact that we do not is a strong indicator that we are digging into the limits of what our strategy can achieve by the end of the second full-bandwidth self-calibration pass. This is likely due to the fact that our initial sky model (and all subsequent models) only include 3C295: the fact that all the other sources in the primary beam are not included in the model is bound to introduce bias our gain solutions and artefacts in our images.

@@

\pg
Because no high-resolution models exist for 3C295 at our observing frequencies, which span quite a large bandwidth, we start by making a model from a high-resolution image taken at 8.7GHz, shown in \cref{fig.vla.3c295}. We start with a model which is as close to the expected ``true" brightness distribution because the self-calibration problem is not convex: while the calibration inverse problem itself is convex (i.e. we expect to converge towards the ``true" gains regardless of the initial conditions, given enough SNR), the degeneracy introduced by imaging in the self-calibration procedure means that one could well get stuck in a local minimum.

\pg
A good initial test of our hypothesis that we can start from a ``good enough" point to reach a global minimum is to self-calibrate on a two-point model. We perform this test, keeping only the two brightest pixels of the model shown above to test whether we converge to the same structure. The result is shown as an overlay in \cref{fig.sc.2pt}: as we can see, we recover the same structure as the full VLA model. We therefore start from this full model for our ``proper" self-calibration procedure.
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/{3C_295_I_8.7GHz_lbs2003.J2000.fits2ptmodel}.png}
\caption{\label{fig.sc.2pt} Overlay of the model resulting from a self-calibration process starting from a two-point model. The overlays start at 0.9$\sigma$ of the final, 60-subband self-cal restored image, and move in increments of 6$\sigma$}.
\end{figure}

\pg
The only exhaustive method to test our hypothesis would be to perform a full Monte Carlo analysis, testing every possible flux distribution for every pixel at every frequency to ensure that we do, in fact, find the model and gains which give the global minimum fit, rather than a local minimum fit. This would be extremely time-consuming, however, and falls outside the scope of this work. We therefore simply try to start as close as possible to where we expect the global minimum to fall, using the full VLA image as a starting model.

\pg
We acquire the initial calibration model by extracting features from a NASA/IPAC Extragalactic Database\footnote{\hyperref[here]{https://ned.ipac.caltech.edu/}} image of 3C295 \citepads[see][]{1991AJ....101.1623P}, shown in Fig. \ref{fig.vla.3c295}. This feature extraction is carried out using PyBDSM \citepads{2015ascl.soft02007M}, changing all extracted Gaussians into points\footnote{This practice ensures that wrongly-estimated Gaussians do not end up introducing unphysical bias during calibration.}.
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/{3C_295_I_8.7GHz_lbs2003.fits}.png}
\caption{\label{fig.vla.3c295}VLA observation of 3C295 at 8.7 GHz. Pixel size is $0.2''$.}
\end{figure}
% command: python ~/Scripts/PngCutoutFromFits.py --filename 3C_295_I_8.7GHz_lbs2003.fits --RA 14:09:33.418 --Dec 52:26:13 --size .12 --vmin -0.5 --vmax 4 --invert

\pg
Because we apply a $uv$-cut of 1km when calibrating, and a $uv$-cut of 100km when imaging, we need to make sure that we do not introduce spurious structure in the core and remote station gains. Similarly, we can check the quality of our self-calibration by looking at the structure of the international station gains. Gain structure can consist of genuine atmospheric \& instrumental effects, but can also include unmodeled flux being absorbed into the gain solutions. At first, this is unavoidable: if flux is resolved by a baseline but not included in the model, then it will be absorbed into the gains. As we improve our model, we expect the gain curves to flatten and become smoother. If this happens for our international station gains, then we have a strong indication that we are in fact recovering physical signal rather than adding artefacts into our model. Similarly, we expect core and remote station gains to remain mostly static: if this is the case, then our strategy does not bias our gain solutions for these stations away from the underlying truth.

\pg
Let us begin by checking whether our core and remote station gains change over time due to our self-calibration procedure. Here, we look at the gains of SB155, on which we applied the initial and final gain solutions. This subband was included in the 60-subband self-calibration step, but not the 6-subband self-calibration step.
\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.firstpass}.png}
\caption{\label{fig.sb155.gains.cs201.before} CS005, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.lastpass}.png}
\caption{\label{fig.sb155.gains.cs201.after} CS005, last calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.RS305.before} RS305, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.RS305.afer} RS305, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.CR} Gain amplitude \& phase for CS201 of SB155}
\end{figure}

\pg
As we can see from \cref{fig.sb155.gains.CR}, there is little change in amplitude structure between the first pass of calibration (with the model extracted from \cref{fig.vla.3c295}) and the last pass of calibration (after the successive self-calibration passes). Similarly, there is little apparent change in phase structure (note that the phase wraps from $-\pi$ to $\pi$). It is important to note, however, that there is a difference in the \textit{level} of the amplitude curve: in the first pass of calibration, the average amplitude value is generally higher than in the last pass of calibration. This is expected behaviour, as there is less diffuse emission in the $uv$-cut imaging model of 3C295 than in the initial calibration model. This is equivalent to applying a choice of weighting scheme which down-weighs diffuse emission (i.e. a form of uniform weighting).

\pg
As for the international station gains, they are shown below for three of the seven international stations used for our observation, in \cref{fig.sb155.gains.de602,fig.sb155.gains.se607,fig.sb155.gains.uk608}. We show only one of the four German stations. As we can see, while structure very clearly remains in some of these amplitude curves (notable in the last two hours, e.g. \cref{fig.sb155.gains.DE602.after}), some of the structure disappears from these amplitude curves. It should be noted that the last two hours of the observation are generally much noisier than the first 6; the last quarter of these plots show this. Quality-based weighting schemes compensate for this effect, and minimise the associated image deterioration. Note that the phase wraps extremely fast in the international station gain plots: this is because phase is calculated relative to CS001HBA0, the first antenna in our array. Because the international stations are much farther away from this core station than other core or remote stations are, we expect the associated phase solutions to wrap faster.
\begin{figure}[h!]
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.DE602.before} DE602, first calibration}
\end{subfigure}
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.DE602.after} DE602, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.de602} Gain amplitude \& phase for DE602, international stations of SB155}
\end{figure}

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.SE607HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.SE607.before} SE607, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.SE607HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.SE607.after} SE607, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.se607} Gain amplitude \& phase for SE607, an international stations of SB155}
\end{figure}

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.UK608HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.UK608.before} UK608, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.UK608HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.UK608.after} UK608, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.uk608} Gain amplitude \& phase for UK608HBA, an international stations of SB155}
\end{figure}

\pg
So far, we have shown the gain solutions for one subband, out of a total of 324 used subbands. The spectral behaviour of the amplitude and phase solutions can give us insight into whether we are absorbing true flux or calibration artefacts into our model: if our calibration model becomes more accurate to the true underlying flux, we expect e.g. the phase of core station calibration solutions to remain the same (as we are improving high-resolution details of our model) while our international station calibration solutions would become flatter in amplitude. Similarly, we expect to see large-scale frequency-dependent behaviour in the phase solutions of international stations, with smaller fluctuations around this behaviour: the effect of differential TEC, ionospheric turbulence, etc. Indeed, while the core stations (and, to a lesser extent, the remote stations) can be approximated as having the ``same ionosphere" above them, this is not the case for international stations: this is a major source of the difficulties in calibrating these stations. As such, for international stations, we expect to see complex structure in the phase gain solutions, characteristic of the differences between the local atmosphere above the core and the international stations. For core stations, we only expect to see structure characteristic of scintillation in the phase gain solutions.

@@

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.gains.CS005.before} CS005HBA0, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.gains.CS005.after} CS005HBA0, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.CS005} Plots showing the amplitude and phase solutions for CS005HBA0 at the start and end of self-calibration, respectively. Some areas are empty: in the case of the first calibration, this is because the solutions were not calculated for all subbands. For the last calibration, this is because some subbands were found to be corrupted, and their solutions are thus not shown here.}
\end{figure}

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS201HBA0.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.CS201HBA0.before} CS201HBA0, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.CS201HBA0.after} CS201HBA0, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.CS201HBA0} Plots showing the amplitude and phase solutions for CS201HBA0 at the start and end of self-calibration, respectively. Some areas are empty: in the case of the first calibration, this is because the solutions were not calculated for all subbands. For the last calibration, this is because some subbands were found to be corrupted, and their solutions are thus not shown here.}
\end{figure}
